Optimizing Structural Properties Of Neural Networks With Genetic Algorithms

### 2015-07-31 paper/paper.tex
introduction extended

\title{Optimizing Structural Properties Of Neural Networks With Evolutionary Algorithms}

@@

\section{Introduction}
What makes artificial neuronal networks learn successfully? And why are
certain types of networks and certain learning rules effective for some
problems, but fail at solving others?
Confronting and eventually answering these questions is critical when trying
to gain a deep understanding of neural networks and artificial
intelligence.

*MISSING*

Instead of fine tuning single internal weights in order to obtain a network
perfectly fit for the task, however, we are concerned with the rules that
generate the structure of the networks. So we still want our networks to be
random to a certain degree, which seems biologically plausible, but we also
want them to follow certain patterns and allow specialization: How dense
are the synapses distributed? Is there a strong compartimization?  Are
recurrent structures frequent? How strong should the feedback signal be,
and how strong should the synapses be on average?

@@

>>>>

\title{Optimizing Structural Properties Of Neural Networks With Genetic Algorithms}

@@

\section{Introduction}
In contemporary artificial neuronal network science a vast number of
different models for describing both the dynamic of networks as well as
their plasticities are applied. 
But what exactly makes artificial neuronal networks learn successfully? And
why are certain types of networks and certain learning rules effective for
some problems, but fail at solving others? Confronting and eventually
answering these questions is critical when trying to gain a deep
understanding of neural networks and artificial intelligence.

Using the above questions as a guideline,
we decided to analyze the FORCE algorithm introduced in \cite{FORCE} that
is especially suitable for learning periodic patterns. More precisely, we
wanted to identify which properties of a network are important for its
learning success, and how to choose these properties' values for an optimal
result.  To this end, one option would be to take an initial network and
an adequate periodic pattern to be learned, and to optimize the network's
internals for this very task. Then patterns may emerge that hint at how an
optimal network design looks like.

However, we took a different approach:
Instead of fine tuning single internal weights, we were concerned with the
general rules of the network's structure. So we wanted our networks to
be random to a certain degree, which seems biologically plausible, but we
also wanted them to follow certain patterns and allow for specialization:
How dense are the synapses distributed? Is there a strong compartmentalization?
Are recurrent structures frequent? How strong should the feedback signal
be, and how strong should the synapses fire on average?

To address these issues we looked at \emph{parameterized random networks}, or
in the language of genetics, at different \emph{network genotypes}. These
genotypes contain parametric information about the networks to be created,
the \emph{phenotypes}. Different phenotypes with the same genotype may thus
have strong quantitative differences, but they share structural qualities
whose fitness for a preassigned task we wanted to analyze and improve.

Since calculating the fitness of a genotype required the evaluation of many
different phenotypes and because of the resulting stochastic nature of the
problems we faced, we decided to use genetic algorithms for the
optimization process. Other more direct optimization methods like simulated
annealing or gradient descent seemed unfeasible for this.

@@

- describe network dynamic
- describe learning and tasks
- describe sucess of network
